Spin canonical divisors of log stacky curves

### 2015-06-24 stacky_write_up/spin_cring.tex
Changes to proofs of section 3 lemmas

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bh{{\mathbb H}}
\newcommand\bO{{\mathbf O}}
\newcommand\CC{{\mathcal C}}
\newcommand\BN{{\mathbb N}}
\newcommand\BC{{\mathbb C}}
\newcommand\BF{{\mathbb F}}
\newcommand\BR{{\mathbb R}}
\newcommand\BQ{{\mathbb Q}}
\newcommand\BP{{\mathbb P}}
\newcommand\BBZ{{\mathbb Z}}
\newcommand\uR{\underline{R}}
\newcommand\uZ{\underline{\BBZ}}
\newcommand\CF{{\mathcal F}}
\newcommand\uCF{\underline{{\mathcal F}}}
\newcommand\BZ{{\mathbb Z}}
\newcommand\BA{{\mathbb A}}
\newcommand\fa{{\mathfrak a}}
\newcommand\fp{{\mathfrak p}}
\newcommand\fq{{\mathfrak q}}
\newcommand\fm{{\mathfrak m}}
\newcommand\so{{\mathscr O}}
\newcommand\sg{{\mathscr G}}
\newcommand \sx{\mathscr X}

@@

Let $\Gamma$ be a {\bf Fuchsian} subgroup, i.e. a discrete subgroup of $PSL_2(\BR)$ acting on the {\bf upper half plane} $\bh$ by fractional linear transformations. We can then consider the orbifold $\Gamma \backslash \bh$ over $\BC$. For example, in the case $\Gamma$ acts freely on $\bh$, $\Gamma \backslash \bh$ is a Riemann Surface over $\BC$. It may well be that $\Gamma \backslash \bh$ is not comapact. Defining $\Delta$ as the divisor of cusps associated to $\Gamma$, we can form a space $\Gamma \backslash \bh^*$ which is a compactification of $\Gamma \backslash \bh,$ with associated {\bf divisor of cusps} $\Delta$. We can now consider the {\bf ring of modular forms} $M(\Gamma) = \oplus_k M_k(\Gamma)$. In this paper, for general $\Gamma$, we sharply bound the degree of the generators and relations for $M(\Gamma).$ This was already answered in \cite[]{vzb:stacky} in the case that $M(\Gamma)$ has $M_{2k}(\Gamma) = 0$ for all $k \in \BZ$, that is, when $\Gamma$ has no odd weight modular forms, and we extend this to all Fuchsian subgroups $\Gamma$.

In order to answer this question, we translate the seemingly analytic question into the algebraic category, using a generalization of the GAGA principle. That is, by \cite[Proposition 6.1.5]{vzb:stacky}, there is an equivalence of categories between orbifold curves and log stacky curves over $\BC$. Hence, for the remainder of the paper, we will work in the algebraic category.

@@

Folloing Voight and Zureick-Brown, \cite{vzb:stacky}, we examine a much more general setting, generalizing Petri's theorem in many directions.
A divisor $\Delta \in \d X$ is a {\bf log divisor} if $\Delta = \sum_{i}^{}P_i \in \di X,$ with all $P_i$ distinct. Let $(\sx,\Delta)$ denote a tame seperably rooted stacky curve over a field $k$ so that $X$ is the coarse space of $\sx$, as defined in \cite{vzb:stacky}[Section 5.2] so that $\di \sx = \div X \bigoplus_{i=1}^r \langle \frac{1}{e_i}P_i \rangle,$ with $e_i \in \BZ$. Let $K_\sx = K_X + \sum_{i=1}^{r}\frac{e_i-1}{e_i}P_i.$ In this case, we say $\sx$ has {\bf stabilizer order} $e_i$ at $P_i$. Call $\halfcan \in \di \sx$ a {\bf log half-canonical divisor} on $(\sx,\Delta)$, be a divisor $\halfcan \in \di \sx$ so that $2\halfcan \sim K_\sx + \Delta.$ If $X$ has genus $g$ and $\delta = \deg \Delta,$ then we say $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$. A {\bf log spin curve} is a triple $(X,\Delta,\halfcan)$ where $(X,\Delta)$ and $\halfcan$ is a log half-canonical divisor on $(\sx,\Delta)$. The central object of study in this paper is the {\bf log spin canonical ring} of $(X,\Delta,L)$ is
\begin{align*}
	R(X,\Delta,L) = \oplus_{n=0}^\infty H^0(X,nL)
\end{align*}
The main theorem of this paper is the following bound on the degree of generators and relations of arbitrary log spin canonical rings:

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has relations in degree at most $\max(11, e_1,
\ldots, e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has relations in degree at most
$\max(10, e_1, \ldots, e_r),$ so long as $\sigma$ does not lie in a finite list of
exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g_0_exceptional}
for signatures with $g = 0$.
\end{thm}
\todo{Say that this follows immediately from the three theorems for the special cases.}
Note that in the course of the proof of Theorem ~\ref{thm:g_0_generators_relations} we will not merely obtain bounds on the degrees of the generators and relations. In fact, if the genus $g = 0$ or $g = 1$, the proof yields an explicit system of generators and the initial ideal of relations. While in the genus $g \geq 2$ case, the proof does not always yield an explicit system of generators and relations, if the halfcanonical divisor $\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',0,X)$ a log spin curve, and a system of generators and relations for $\halfcan'$ is known, then our proof provides an inductive procedure for determining the generators and initial ideal of relations for $\halfcan$. In fact, such a system of generators and relations are known for $\halfcan'$ in many cases, as detailed in Neves ~\cite[Section III.4]{neves:halfcan}.

@@

\section{Background}
\label{sec:background}
\todo{Reference independence of base field}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
\todo{edit this}

@@

\begin{defn}
\label{defn:sat}
\textbf{Saturations} of the canonical divisor, as defined in Voight
and Zureick-Brown ~\cite[Section 7.2]{vzb:stacky}.
\end{defn}

@@

\begin{lem}
\label{lem:deg1_sat_ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper, or preferably define it in the background section})
satisfying $\deg(\lfloor{D}\rfloor)\ge 0$.  Suppose $P$ is not a base point of $D$ meaning that $R_D$ has a
generator $u$ in degree 1 that is a section of $\so_D$ with $\ord_P(u)=0$.  Let $D$ be generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
in degree at most $\bar{\tau}$.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbf{N}$ such
that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \forall d \in \mathbf{
	N} : d \ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

@@

\begin{lem}
\label{lem:sat-two-induction-high-genus}
Let $(\sx, \Delta, \halfcan_\sx)$ and $(\sx', \Delta, \halfcan_{\sx'})$ be log stacky curves with coarse space $(X, \Delta, L_X)$ and signatures $(g; e_1, \ldots, e_r; \delta)$ and $(g, e_1, \ldots, e_{r-1}, \delta)$ respectively, with $g > 0,$ and, if $g = 1$ then $\deg 3L \geq 2$.
Let $R_{L'} = R' = k[x]/I'= k[x_2, x_3
, x_5,\ldots, x_m]/I'$ and let $\halfcan = \halfcan' + \frac{1}{
3}P$, where $P\in X$ is a base point of $\halfcan'$ (which includes the case when $H^0(X,\lfloor \halfcan'\rfloor)=0$).  Suppose $\halfcan$ and $\halfcan'$ satisfy $sat(Eff(\halfcan')) = sat(Eff(\halfcan)) = 2.$
Suppose $\deg x_i = i$ for $i \in \{2, 3\}$ and choose an ordering on $k[x_2, \ldots, x_m]$ satisfies
\begin{align*}
	\ord_{x_2}(f) < \ord_{x_2}(g) \implies f \prec g.
\end{align*}
Then, the following statements hold.
\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3, 4\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4]$ with $grlex$ so that $y_4 \prec 
		y_3$
		and equip the ring $k[y_3, y_4, x]$ with the block 
		order so that $R = k[y_3, y_4, x]/I$. Let,
		\begin{align*}
			\initial_\prec(I) &= \initial_\prec(I')k[x, y_3, y_4] \\
			&+\langle y_4 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_4^2.\rangle
		\end{align*}
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{lem}

\begin{proof}
The assumptions on $g,\sigma$ imply $H^0(\sx, 3L), H^0(\sx, 4L)$ are both base point free: If $g \geq 2$ then $\deg 3L > 2g-1$ and $\deg 4L > 2g-1$, so $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. If $g = 1$, we assume $\deg 3L \geq 2 > 2g-1$, so we also have $\deg 4L \geq 2 > 2g -1,$ so again $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. Therefore, general elements $y_3$ and $y_4$ satisfy $-\ord
_P(y_i) = 1$ by Riemann Roch, and the assumptions on $\sigma$. \todo{Understand why/whether we need 
to cite lemma 5.4.7.}

@@

To show part (b), note that the generators in
~\ref{eqn:sat_two_add_generator} are precisely a set of monomials which
generate $k[x,y]/J$ over $k$. So, to show $J = in_\prec(I)$, it
suffices to show that all generators of $J$ lie in $in_\prec(I)$.

This follows, since there exist constants $A_j \in k$ for $3 \leq j \leq m,$ and a constnant $B \in k$ and
elements $w_{i,j} \in R'$ so that the following linear combinations
of elements lie in $R'$.

\begin{align*}
	y_4x_j &- A_j y_3 w_j & \text{ so that } 2 \leq j \leq m \text{ and } \deg w_j = \deg x_j + 1\\
	y_4^2 &+ B y_3^2 x_2 \\
\end{align*}

>>>>

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bO{{\mathbf O}}
\newcommand\CC{{\mathcal C}}
\newcommand\BN{{\mathbb N}}
\newcommand\BC{{\mathbb C}}
\newcommand\BF{{\mathbb F}}
\newcommand\BR{{\mathbb R}}
\newcommand\BQ{{\mathbb Q}}
\newcommand\BP{{\mathbb P}}
\newcommand\BBZ{{\mathbb Z}}
\newcommand\uR{\underline{R}}
\newcommand\uZ{\underline{\BBZ}}
\newcommand\CF{{\mathcal F}}
\newcommand\uCF{\underline{{\mathcal F}}}
\newcommand\BZ{{\mathbb Z}}
\newcommand\BA{{\mathbb A}}
\newcommand\fa{{\mathfrak a}}
\newcommand\fp{{\mathfrak p}}
\newcommand\fq{{\mathfrak q}}
\newcommand\fm{{\mathfrak m}}
\newcommand\so{{\mathscr O}}
\newcommand\sg{{\mathscr G}}
\newcommand \sx{\mathscr X}

@@

Example, stuff from Voight and Zureick-Brown \cite{vzb:stacky}
and O'Dorney's paper \cite{dorney:canonical}.

We consider divisors of the form $D = \sum_{i = 1}^{r} \subhalf{e_i}$.

\subsection{Main Results}
\begin{enumerate}
	\item Modular Forms
	\todo{Reference independence of base field}
\end{enumerate}

@@

In Voight and Zureick-Brown, \cite{vzb:stacky} examine a much more general setting, generalizing Petri's theorem in many directions.
A divisor $\Delta \in \d X$ is a {\bf log divisor} if $\Delta = \sum_{i}^{}P_i \in \di X,$ with all $P_i$ distinct. Let $(\sx,\Delta)$ denote a tame seperably rooted stacky curve over a field $k$ so that $X$ is the coarse space of $\sx$, as defined in \cite{vzb:stacky}[Section 5.2] so that $\di \sx = \div X \bigoplus_{i=1}^r \langle \frac{1}{e_i}P_i \rangle,$ with $e_i \in \BZ$. Let $K_\sx = K_X + \sum_{i=1}^{r}\frac{e_i-1}{e_i}P_i.$ In this case, we say $\sx$ has {\bf stabilizer order} $e_i$ at $P_i$. Call $\halfcan \in \di \sx$ a {\bf log half-canonical divisor} on $(\sx,\Delta)$, be a divisor $\halfcan \in \di \sx$ so that $2\halfcan \sim K_\sx + \Delta.$ If $X$ has genus $g$ and $\delta = \deg \Delta,$ then we say $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$. A {\bf log spin curve} is a triple $(X,\Delta,\halfcan)$ where $(X,\Delta)$ and $\halfcan$ is a log half-canonical divisor on $(\sx,\Delta)$. The central object of study in this paper is the {\bf log spin canonical ring} of $(X,\Delta,L)$ is
\begin{align*}
	R(X,\Delta,L) = \oplus_{n=0}^\infty H^0(X,nL)
\end{align*}
The main theorem of this paper is the following bound on the degree of generators and relations of arbitrary log spin canonical rings:

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has relations in degree at most $\max(11, e_1,
\ldots, e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has relations in degree at most
$\max(10, e_1, \ldots, e_r),$ so long as $\sigma$ does not lie in a finite list of
exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g_0_exceptional}
for signatures with $g = 0$.
\end{thm}
\todo{Say that this follows immediately from the three theorems for the special cases.}
Note that in the course of the proof of Theorem ~\ref{thm:g_0_generators_relations} we will not merely obtain bounds on the degrees of the generators and relations. In fact, if the genus $g = 0$ or $g = 1$, the proof yields an explicit system of generators and the initial ideal of relations. While in the genus $g \geq 2$ case, the proof does not always yield an explicit system of generators and relations, if the halfcanonical divisor $\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',0,X)$ a log spin curve, and a system of generators and relations for $\halfcan'$ is known, then our proof provides an inductive procedure for determining the generators and initial ideal of relations for $\halfcan$. In fact, such a system of generators and relations are known for $\halfcan'$ in many cases, as detailed in Neves ~\cite{neves:halfcan}[Section III.4].

@@

\section{Background}
\label{sec:background}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
\todo{edit this}

@@

\begin{defn}
\label{defn:sat}
\textbf{Saturations} of the canonical divisor, as defined in 7.
2 of VZB \cite{vzb:stacky}.
\end{defn}

@@

\todo{definition and basic properties of best lower approximations}

\begin{lem}
\label{lem:deg1_sat_ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper, or preferably define it in the background section})
satisfying $h^0(X, \lfloor{D}\rfloor)\ge 1$.  Suppose $P$ is not a base-point of $dD$ for all $d\in \mathbf{N}^+$ meaning that $R_D$ has generically chosen generators (in the appropriate respective degrees)
$u, x_1, \ldots, _m$ all with $\ord_P=0$ and such that $\deg(u)=1$.  Let $D$ be generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
in degree at most $\bar{\tau}$, where all $x_i$s also satisfy $\ord_P(u)(x_i)$.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbf{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \; \forall d \in \mathbf{
	N} : d \ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

@@

\begin{lem}
\label{lem:sat-two-induction-high-genus}
Let $(\sx, \Delta, \halfcan_\sx)$ and $(\sx', \Delta, \halfcan_{\sx'})$ be log stacky curves with coarse space $(X, \Delta, L_X)$ and signatures $(g; e_1, \ldots, e_r; \delta)$ and $(g, e_1, \ldots, e_{r-1}, \delta)$ respectively, with $g > 0,$ and, if $g = 1$ then $\deg 3L \geq 2$.
Let $R_{L'} = R' = k[x]/I'= k[x_2, x_3
, x_5,\ldots, x_m]/I'$ and let $\halfcan = \halfcan' + \frac{1}{
3}P$, where $P\in X$ is a base point of $\halfcan'$ (which includes the case when $H^0(X,\lfloor \halfcan'\rfloor)=0$).  
Suppose for $i \in \{2, 3\}$, $\deg x_i = i$ and $\ord_P(x_i)=0$ (which is satisfied generically in degree 2 and 3, since base-point freeness follows by Riemann-Roch).  Choose an ordering on $k[x_2, \ldots, x_m]$ that satisfies
\begin{align*}
	\ord_{x_2}(f) < \ord_{x_2}(g) \implies f \prec g.
\end{align*}
Then, the following statements hold.
\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3, 4\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4]$ with $grlex$ so that $y_4 \prec 
		y_3$
		and equip the ring $k[y_3, y_4, x]$ with the block 
		order so that $R = k[y_3, y_4, x]/I$. Let,
		\begin{align*}
			\initial_\prec(I) &= \initial_\prec(I')k[x, y_3, y_4] \\
			&+\langle y_4 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_4^2.\rangle
		\end{align*}
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{lem}

\begin{proof}
The assumptions on $g,\sigma$ imply $H^0(\sx, 3L), H^0(\sx, 4L)$ are both base point free: If $g \geq 2$ then $\deg 3L > 2g-1$ and $\deg 4L > 2g-1$, so $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. If $g = 1$, we assume $\deg 3L \geq 2 > 2g-1$, so we also have $\deg 4L \geq 2 > 2g -1,$ so again $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. Therefore, general elements $y_3$ and $y_4$ satisfy $-\ord
_P(y_i) = 1$ by Riemann-Roch, and the assumptions on $\sigma$. \todo{Understand why/whether we need to cite lemma 5.4.7.}

@@

For any $x_j$, notice that $-\ord_P(y_4x_j)\le 1$.  If it is nonpositive, then it lies in $R_{\halfcan'}$ immediately giving us a relation with initial term $y_4x_j$.  Otherwise, 
we can choose $A_j \in k$ and $w_j\in R_{\halfcan'}$ such that $y_4x_j-A_jy_3w_j$ has no pole at $P$ and thus lies in $R_{\halfcan'}$ giving us a desired relation.  
Similarly, we can find $B, C \in k$ such that $y_4^2 + B y_3^2 x_2+ Cy_3x_2x_3$ has no pole at $P$ and is thus in $R_D$.  In both cases we induce a relation, with initial term described in the lemma's statement.  Noting that any monomial is either a multiple of one of these initial terms, is a basis element (described above), or is in $R_D$, we see that these monomials in fact generate the initial ideal of $I$ over $\initial(I')$.
